---
title: "Homework 5"
author: "Gianni Spiga"
date: '2022-10-30'
output: html_document
---

```{r, message=FALSE}
library(ggplot2)
library(GGally)
library(plotly)
```
## Question 2 

d.) 
```{r}
Xh <- c(1, 0, 0, 0)
XtXinv <-
  matrix(rbind(
    c(0.087,-0.014,-0.035,-0.004),
    c(-0.014, 0.115,-0.012,-0.052),
    c(-0.035,-0.012, 0.057,-0.014),
    c(-0.004,-0.052,-0.014, 0.050)
  ), nrow = 4, ncol = 4)
XtXinv


SE.predh <- sqrt(1.040 * (1 + t(Xh) %*% XtXinv %*% Xh))
SE.predh

qt(1 - (0.05 / 2), 30 - 4)

c(0.9918 - (qt(1 - (0.05 / 2), 30 - 4) * SE.predh), 0.9918 + (qt(1 - (0.05 / 2), 30 - 4) * SE.predh))
```

## Question 6

a.)
```{r}
property <- read.table("property.txt")
colnames(property) <-
  c("Ren.Rate", "Age", "Exp", "Vac.Rate", "Sq.Foot")
property
```

For the variable types, Age, Vacancy Rate, and Total Square Footage are type doubles, Operating Expenses and Rental Rate are integers. Now we observe the distribution of each variable and some summary statistics. 

```{r}
ggplot(data = property, aes(x = Ren.Rate)) + geom_histogram(bins = 10, fill = "#55A3C0")
summary(property$Ren.Rate)
```

We can see approximate symmetry with light tails in the age variable. The mean and the median are very close to each other. 

```{r}
ggplot(data = property, aes(x = Age)) + geom_histogram(bins = 10, fill = "#556EC0")
summary(property$Age)
```

For the Age variable, we can see a bimodal distribution. The majority of the data sits between 0 and 5 thousand and 10 and 20 thousand. Because of this spread, the median is lower, towards where a larger bulk of the data is. The mean however is leaning towards the center of the data, being skewed by the concentration of larger values. 

```{r}
ggplot(data = property, aes(x = Exp)) + geom_histogram(bins = 13, fill = "#7255C0")
summary(property$Exp)
```

With our Opertating Expenses variable, we observe a left skewed distribution. Most of the data is within our IQR, from about 8.13 to 11.62. 

```{r}
ggplot(data = property, aes(x = Vac.Rate)) + geom_histogram(bins = 9, fill = "#7255C0")
summary(property$Vac.Rate)
```

In the Vacancy Rate data, we can see a very strong right skew. Most of our data does not exceed 0.1. Because of the skewing, we can see that the median is smaller than the mean. 

```{r}
ggplot(data = property, aes(x = Sq.Foot)) + geom_histogram(bins = 7, fill = "#7255C0")
summary(property$Sq.Foot)
```

With our Square Footage variable, we can again see a right skew. Unlike Vacancy Rate, this skew is not as strong as it has some pull towards the center, as well as a heavier tail. 

b.) 
```{r, warning = FALSE}
ggplotly(ggpairs(property, title = "Correlogram of Property Data"))
```

From the plot, we can see that the variables with the highest correlation is Rental Rate and Square Footage. The variables with the smallest correlation are Rental Rate and Vacancy Rate. This makes sense, the scatter plot for Rental Rate and Square Footage has the most linear resemblence while Rental Rate and Vacancy Rate show almost no relationship. We can see that Vacancy Rate is negatively correlated with Age and Operating Expenses. 

c.)

```{r}
model1 <- lm(Ren.Rate ~ Age + Exp + Vac.Rate + Sq.Foot, data = property)
summary(model1)
```

The least squares estimates are the following:
```{r}
model1$coefficients
```
The fitted regression function: 
$$
Y_i = 12.220 - 0.142 X_{i1} + 0.282X_{i2} + 0.619X_{i3} + 7.924e-06X_{i4}
$$

The $R^2$ for the model is $0.5847$ and the $R^2_{adj}$ is $0.5629$. 

We can print the ANOVA table to gather more information.

```{r}
#MSE
anova(model1)
```
The ANOVA table shows us that our $MSE = 1.293$. 

d.) 
```{r}
#Residuals vs Fitted Values
ggplot() + aes(x = model1$fitted.values, model1$residuals) + geom_point() + labs(x = "Fitted Values", y = "Residuals", title = "Residuals vs. Fitted Values") + geom_abline(intercept = 0, slope = 0)

## Normal Q-Q plot
ggplot(model1, aes(sample = model1$residuals)) + stat_qq() + stat_qq_line() + labs(x = "Theoretical Quantiles", y = "Sample Quantiles", title = "Normal Q-Q")

# Residual Boxplot
ggplot() + aes(x = model1$residuals) + geom_boxplot() + labs(x = "Residuals") 
```

We can see from the Residual vs Fitted Value plot that we have random scatter with no sign of a pattern, suggesting that a linear model is the best model for this data. Our Normal QQ Plot shows us that our distribution is approximately normal with signs of light tails. Our box plot shows us that our data is approximatly symmetric, with only 4 outliers in the data set. 

```{r}
#Residuals vs Age
ggplot() + aes(x = property$Age, y = model1$residuals) + geom_point() + labs(x = "Age", y = "Residuals", title = "Residuals vs. Age") + geom_abline(intercept = 0, slope = 0)

# Residuals vs Operating Expenses
ggplot() + aes(x = property$Exp, y = model1$residuals) + geom_point() + labs(x = "Operating Expenses", y = "Residuals", title = "Residuals vs. Operating Expenses") + geom_abline(intercept = 0, slope = 0)

# Residuals vs Operating Expenses
ggplot() + aes(x = property$Vac.Rate, y = model1$residuals) + geom_point() + labs(x = "Vacancy Rate", y = "Residuals", title = "Residuals vs. Vacancy Rate") + geom_abline(intercept = 0, slope = 0)

# Residuals vs Square Footage
ggplot() + aes(x = property$Sq.Foot, y = model1$residuals) + geom_point() + labs(x = "Square Footage", y = "Residuals", title = "Residuals vs. Square Footage") + geom_abline(intercept = 0, slope = 0)


### Interactions 

#Residuals vs Age*Expenses
ggplot() + aes(x = property$Age * property$Exp, y = model1$residuals) + geom_point() + labs(x = "Age*Expenses", y = "Residuals", title = "Residuals vs. Age*Expenses") + geom_abline(intercept = 0, slope = 0)

#Residuals vs Age*Vac.Rate
ggplot() + aes(x = property$Age * property$Vac.Rate, y = model1$residuals) + geom_point() + labs(x = "Age*Vac.Rate", y = "Residuals", title = "Residuals vs. Age*Vac.Rate") + geom_abline(intercept = 0, slope = 0)

#Residuals vs Age*Sq.Foot
ggplot() + aes(x = property$Age * property$Sq.Foot, y = model1$residuals) + geom_point() + labs(x = "Age*Sq.Foot", y = "Residuals", title = "Residuals vs. Age*Sq.Foot") + geom_abline(intercept = 0, slope = 0)

#Residuals vs Expenses*VacRate
ggplot() + aes(x = property$Exp * property$Vac.Rate, y = model1$residuals) + geom_point() + labs(x = "Exp*Vac.Rate", y = "Residuals", title = "Residuals vs. Exp*Vac.Rate") + geom_abline(intercept = 0, slope = 0)

#Residuals vs Expenses*Sq.Foot
ggplot() + aes(x = property$Exp * property$Sq.Foot, y = model1$residuals) + geom_point() + labs(x = "Exp*Sq.Foot", y = "Residuals", title = "Residuals vs. Exp*Sq.Foot") + geom_abline(intercept = 0, slope = 0)

#Residuals vs Vac.Rate*Sq.Foot
ggplot() + aes(x = property$Vac.Rate * property$Sq.Foot, y = model1$residuals) + geom_point() + labs(x = "Vac.Rate*Sq.Foot", y = "Residuals", title = "Residuals vs. Vac.Rate*Sq.Foot") + geom_abline(intercept = 0, slope = 0)
```

We have six interaction terms. When we observe plots with interactions with Vacancy Rate, we notice that a lot of the observations are very close to zero. Because of this, the data does not have random scatter, but instead, high concentration to the left with outliers on the right. We see this same scenario with interactions with Total Square Footage variable, however not as strong. 

In our non-interaction plots, we can see that variables Age and Vacancy Rate have clusters of data, which violates our desire for random scatter. 

f.) To test our Beta coefficients, we have: 
$$
H_0: \beta_k = 0 \\
H_a: \beta_k \neq 0
$$
With test statistic and null distribution:
$$
T^* = \frac{\hat{\beta_k}}{s(\hat{\beta_k})} \\
T^* \sim t_{n-p}
$$

We reject $H_0$ if $T^* > t(1 - \alpha/2, n-p)$.

Thankfully, the summary of our model does this test for us with the p-values. For 