---
title: "Homework 2"
author: "Gianni Spiga"
date: '2022-10-11'
output: html_document
---

## Question 6
a.) The relationship betweein high school graduates and crime rate, according to the trendline in the scatterplot, has a moderate strength, negative relationship. As the percentage of graduates increases, the crime rate decreases. 

b.)
```{r}
n <- 84
Xi <- 6602
Yi <- 597341
Xi2 <- 522098
Yi2 <- 4796548849
XiYi <- 46400230

Xbar <- Xi/n
Ybar <- Yi /n

Beta1 <- (XiYi - (n*Xbar*Ybar)) / (Xi2 - n*(Xbar)^2)
Beta1

Beta0 <- Ybar - Beta1*Xbar
Beta0
```

Observing our $\beta_0$, we can understand that by our model, if we were to have a county with a high school graduate percentage of 0, we would expect to see a 20,518 crimes per 100,000 people. Our $\beta_1$ tells us that for each percentage increase in high school graduates, we see crimes per 100,000 people drop by on average 171 people. 


$$
SSE = \sum Y_i^2 - 2 \hat{\beta_1} * \sum Y_i - 2 \hat{\beta_1}\sum X_iY_i + n\hat{\beta_0} + 2\ \hat{\beta_0} \hat{\beta_1} \sum X_i + \hat{\beta_1}^2\sum X_i^2
$$
c.)
```{r}
SSE <-
  Yi2 - 2 * Beta0 * Yi - 2 * Beta1 * XiYi + Beta0 ^ 2 * 84 + 2 * Beta0 *
  Beta1 * Xi + Beta1 ^ 2 * Xi2
SSE

MSE <- SSE / (84 - 2)
MSE
```

The degrees of freedom for the SSE is $n - 2$ which in this case is 82. 

d.)
```{r}
# Standard Error of Beta0
seBeta0 <- sqrt(MSE * ((1/n) + (Xbar^2/(Xi2 - n*(Xbar)^2))))
seBeta0

# Standard Error of Beta1
seBeta1 <- sqrt(MSE / ((Xi2 - n*(Xbar)^2)))
seBeta1

Sxx <-((Xi2 - n*(Xbar)^2))
```

e.)
We are testing
$$
H_0: \beta_1 = 0\\
H_a: \beta_1 \neq 0
$$
With test statistic:
$$
T^* = \frac{\hat{\beta_1}} {s(\hat{\beta_1})}
$$
where..
$$
T^* \sim t_{(n-2)}
$$

```{r}
# We test for Beta 1 

Tstar <-  Beta1 / seBeta1
Tstar

critval <- qt(1 - (0.01/2), n - 2)
critval

pval <-pt(Tstar, n -2)
pval
```
Since our critical value is less than our $T^*$, we reject $H_0$, concluding there is a significant relationship between the percentage of high school graduates in a county and that county's respective crime rate per 100,000 people. 

f.) An unbiased estimator for the population parameter $\beta_0$ is our sample intercept $\hat{\beta_0}$.
```{r}
#We create a confidence interval for Beta0
c(Beta0 - qt(1 - 0.01 / 2, n - 2, ) * seBeta0,
  Beta0 + qt(1 - 0.01 / 2, n - 2, ) * seBeta0)

```
We are 99% confident that the true value of $\beta_0$ is contained within the interval $[11874.05, 29161.15]$. 

g.)
```{r}
Yh <- Beta0 + Beta1 * 85
seYh <- sqrt(MSE * (1/928 + (85 - Xbar)^2 / Sxx ))

c(Yh - qt(1 - 0.05/2, n - 2) * seYh, Yh + qt(1 - 0.05/2, n - 2) * seYh)

```
We are 95% confident that the true value of the mean crime rate for counties with a high school graduation rate of 85% is contained within the interval $[5467.109, 6570.308]$. 

h.)
```{r}
Yh <- Beta0 + Beta1 * 85
seYhpred <- sqrt(MSE * (1 + 1/928 + (85 - Xbar)^2 / Sxx))

c(Yh - qt(1 - 0.05/2, n - 2) * seYhpred, Yh + qt(1 - 0.05/2, n - 2) * seYhpred)
```
We prediction with 95% confident that the true value of the crime rate for County A is contained within the interval $[1298.95, 10738.47]$. We can see that this interval is larger than our interval for our mean response in part g, since we have to account for the uncorrelated errors associated with our observations. 

i.) Parts (e) - (h) rely the assumption that $Y_H$ is normally distributed, this way we can utilize the t statistic in both our confidence and prediction intervals.We are also assuming our linear estimators $\beta_0$ and $\beta_1$ follow normal distributions for the same utilization.  

## Question 7

a.)
```{r}
SSR <- Beta1^2 * Sxx
SSTO <- SSE + SSR

SSE
SSR
SSTO

```

b.)
```{r}
# Mean Squares
MSR <- SSR / 1 

MSE
MSR
```

c.)
```{r}
ANOVAtable <-
  data.frame(
    c(SSR, SSE, SSTO),
    c(1, n - 2, n - 1),
    c(MSR, MSE, NA),
    c(MSR / MSE, NA, NA),
    row.names = c("Regression", "Error", "Total")
  )
names(ANOVAtable) <- c("SS", "d.f.", "MS", "F*")
ANOVAtable
```

d.)
$$
H_0: \beta_1 = 0\\
H_a: \beta_1 \neq 0
$$
With test statistic:
$$
F^* = \frac{MSE} {MSE}
$$
where the null distribution is...
$$
F^* \sim F_{(1 - \alpha, n-2)}
$$
and decision rule of rejectiong $H_0$ when...

$$
F^* > F_{(1 - \alpha, 1, n-2)}
$$
```{r}
Fstat <- MSR/MSE
critval <- qf(1 - 0.01, 1, n-2)

critval
Fstat > critval
```

Since our $F^*$ is greater than our critical value, we reject $H_0$, concluding there is a linear association between crime rate and percentage of high school graduates. 

e.) We notice that both our F test and T test provide us the same conclusion, that there is a linear association between crime rate and percentage of high school graduates. There is also a direct relationship that our F statistic is the same as the square of our T statistic. 

f.)
```{r}
R2 <- SSR/SSTO
R2
```

We can see that about 17.03% of the variation in crime rate for a county can be explained by the variation in the percentage of high school graduates. 

```{r}
r <- (XiYi - (n*Xbar*Ybar)) / sqrt(Sxx * (Yi2 - n*(Ybar)^2))
r

r^2
```

We notice that our $R^2$ and the square of our $r$ value are the same value. 