---
title: "Homework 3"
author: "Gianni Spiga"
date: '2022-10-15'
output: html_document
---

## Question 1

a.)
```{r}
library(ggplot2)
library(plotly)
```

```{r}
muscle <- read.table("muscle.txt")
colnames(muscle) <- c("mass", "age")
head(muscle, 10)

ggplotly(ggplot(muscle, aes(x = mass)) + geom_histogram(bins =15, fill = "#FF6347"))
ggplotly(ggplot(muscle, aes(x = age)) + geom_histogram(bins = 15, fill = "#FFA07A"))
```
We can see from the plots that mass has a slight skew to the right while the age histrogram has a strong right skew.

```{r}
ggplotly(ggplot(muscle, aes(x = age, y= mass)) + geom_point(color = "#FFC300"))
```

We can see that from the scatter plot that the data seems to have a negative linear relationship. 

b.)
```{r}
library(MASS)
bxcx <- boxcox(mass ~ age, data = muscle)
lambda <- bxcx$x[which.max(bxcx$y)]
lambda
```

Since our Box Cox 95% confidence interval contains 1, we can conclude that a transformation of our variable is unneccessary. 

c.)  
```{r}
#Build the model
modelMM <- lm(mass ~ age, data = muscle)
sumModelMM <- summary(modelMM)

# Coefficients and their standard errors
sumModelMM$coefficients

# MSE
MSE <- sum(sumModelMM$residuals^2) / (nrow(muscle) - 2)
MSE

# Degrees of Freedom for MSE
sumModelMM$df[2]
```

d.) The regression line for the data is the following:
$$
\hat{Y_i} = 156.347 - 1.190X_i
$$

```{r}



```