---
title: "Homework 7"
author: "Gianni Spiga"
date: '2022-11-17'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    df_print: paged
---

## Question 2

### a.) 
```{r}
cars <- read.csv("Cars.csv")
cars
```
### b.)
```{r}
# Fixing horsepower column
cars$horsepower <- as.numeric(cars$horsepower)

# Find index of NA rows
NArows <- which(is.na(cars))

# View rows
cars[NArows,]
```

### c.) 
```{r}
str(cars)
```

We should treat the variables "cylinders" and "country.code" as categorical variables. The rest of the variables should be treated as quantitative variables. 

```{r}
cars$cylinders <- as.factor(cars$cylinders)
cars$country.code <- as.factor(cars$country.code)
```

### d.) 
```{r}
library(ggplot2)
# MPG
ggplot(data = cars, aes(x = mpg)) + geom_histogram(bins = 15)

#Displacement
ggplot(data = cars, aes(x = displacement)) + geom_histogram(bins = 10)

# Horsepower
ggplot(data = cars, aes(x = horsepower))+ geom_histogram(bins = 20)

# Weight
ggplot(data = cars, aes(x = weight)) + geom_histogram()

# Acceleration
ggplot(data = cars, aes(x = acceleration)) + geom_histogram()
```

From the histograms we notice the following: is slightly right skewed, displacement is heavily right skewed, horsepower and weight is moderately right skewed, and acceleration is approximately normal. 