---
title: "Homework 4"
author: "Gianni Spiga"
date: "2022-10-26"
output: html_document
---
## Question 6

a.) 
```{r}
X <- cbind(rep(1,5), c(-0.63, 0.18, -0.84, 1.6, 0.33), c(-0.82, 0.49, 0.74, 0.58, -0.31))
Y <- c(-0.97, 2.51, -0.19, 6.53, 1)

# X'X 
t(X) %*% X

#X'Y
t(X) %*% Y

# (X'X)^-1 
solve(t(X) %*% X)
```

b.) 
```{r}
B <- solve(t(X) %*% X) %*% t(X) %*% Y
B
```

c.) 
```{r}
H <- X %*% solve(t(X) %*% X) %*% t(X)
H
```
The rank of $H$ is $p = 3$, and the rank of $I_n - H$ is $5-3 = 2$

d.)
```{r}
sum(diag(H))
```
We find that the trace of H is equal to the rank of H, in this case, 3. 

e.) 
```{r}
# Fitted values
Yhat <- X %*% B
Yhat

# Residuals
e <- Y - Yhat
e

#SSE
SSE <- sum((Y - Yhat)^2)
SSE

# MSE
MSE <- SSE / (5 - 3)
MSE
```
The degrees of freedom for the SSE is $n - p = 5 - 3 = 2$.

f.) Now we create the interaction model:
```{r}
Xint <- cbind(X, X[,2] * X[,3])
Xint

Hint <- Xint %*% solve(t(Xint) %*% Xint) %*% t(Xint)
Hint
```

The rank of H is $p = 4$. The rank of $I_n - H = n -p = 5 - 4 = 1$. We notice that the rank of $H$ increases and the rank of $I_n - H$ decreases with the interaction model. 

g.)
```{r}
B.int <- solve(t(Xint) %*% Xint) %*% t(Xint) %*% Y
B.int
```
h.) 
```{r}
Yhat.int <- Xint %*% B.int
Yhat.int

# Residuals
e.int <- Y - Yhat.int
e.int

#SSE
SSE.int <- sum((Y - Yhat.int)^2)
SSE.int

# MSE
MSE.int <- SSE.int / (5 - 4)
MSE.int
```
The degrees of freedom for the SSE is $n -p = 5 - 4 = 1$. 

i.) The model with the interaction terms seems to fit the data better, since we have much smaller residuals and a smaller MSE. 


