---
title: "Homework 6"
author: "Gianni Spiga"
date: '2022-11-08'
output: 
  prettydoc::html_pretty:
    theme: architect
    toc: TRUE
    df_print: paged
---

## Question 1 

### a.) 
True, if the response variable has correlation 0 with all the predictor variables, then the only predictor would be the intercept, with slope 0. In the simple linear regression case, this would be a horizontal line going through the data

### b.) 

### e.) 
True, we can have a large amount of variables that are uncorrelated with each other, which individually can be significant but create a not significant p-value as a whole. 


```{r, message = FALSE}
library(ggplot2)
library(GGally)
library(plotly)
```


```{r}
property <- read.table("property.txt")
colnames(property) <-
  c("Ren.Rate", "Age", "Exp", "Vac.Rate", "Sq.Foot")
property
```

## Question 8 

### a.) 
```{r}
ggplotly(ggplot(data = property, aes(x = Age, y = Ren.Rate)) + geom_point())
```

We can see from the plot that there is no tell of a linear relationship between the age of a property and it's rental rate. 

### b.) 
We have model equation: 
$$
Y_i = \beta_0 + \beta_1\tilde{X_{i1}} + \beta_2X_{i2} + \beta_4X_{i4} +  \beta_1\tilde{X_{i1}^2}
$$
NOTE: can fit X1 for X1 tilde

```{r}
property["AgeCent"] <- property$Age - mean(property$Age)
property["AgeSq"] <- property$AgeCent ^ 2

polyModel <-
  lm(Ren.Rate ~ AgeCent + AgeSq + Exp + Sq.Foot, data = property)


#Plotting Observations Against Fitted Values
ggplotly(
  ggplot() + aes(x = polyModel$fitted.values, y = property$Ren.Rate) + geom_point() + labs(x = "Fitted Values", y = "Observations", title = "Observartions against Fitted Values")
)
```


